---
title: "Increased risks of mortality in the 21st century (2001-2020) attributable
  to PM₂.₅ in Australian Great Capital Cities"
output:
  html_document: default
  word_document: default
  pdf_document: default
---
Lucas Hertzog<sup>1,2</sup>, Edward Jegasothy<sup>3</sup>, Karthik Gopi<sup>3,4</sup>, Cassandra Yuen<sup>1,3</sup>, Dana Jordan<sup>3</sup>, Geoffrey Morgan<sup>3,4</sup>, Ivan Hanigan<sup>1,2</sup>

1. School of Population Health, Faculty of Health Sciences, Curtin University, WA 6102, Australia
2. WHO Collaborating Centre for Climate Change and Health Impact Assessment, WA 6102, Australia
3. School of Public Health, Faculty of Medicine and Health, University of Sydney, Camperdown, NSW 2006, Australia 
4. University Centre for Rural Health, Faculty of Medicine and Health, University of Sydney, Lismore, NSW 2480, Australia

## Outcome
### Average Daily Mortality 2001-2020 (AIHW all causes and ages)

```{r, echo=FALSE}
aihw <- "figures_and_tables/Fig1_mort_avg_doy_all.png"
knitr::include_graphics(aihw)

```

## Exposure
### PM₂.₅ (µg/m³) time-series 2001-2020
```{r, echo=FALSE}

p <- "figures_and_tables/fig_pm25_ts_gcc.png"

# Include the plot in the report
knitr::include_graphics(p)
```

## Results

### Assumptions
PM₂.₅ Predicted = Daily PM2.5 predictions from the random forest model (Bushfire_specific_PM25_Aus_2001_2020_v1_3).

The STL method is used to decompose the time-series into three components: trend, seasonal, and remainder.

The counterfactual is calculated as the sum of the seasonal and trend components, i.e., Counterfactual = Seasonal + Trend.

Following the HRAPIE project's recommendations for concentration–response functions for cost–benefit analysis of particulate matter, the estimated Relative Risk (RR) per 10 μg/m³ increase in PM₂.₅ is 1.0123, with a 95% Confidence Interval (CI) of 1.0045 to 1.0201:

$$
RR_{10\, \mu g/m^3} = 1.0123 \, (95\% \, CI: 1.0045 - 1.0201)
$$
Then, we calculate beta:

$$
\beta = \frac{\log(RR_{10\, \mu g/m^3})}{10}
$$

Subsequently, we define the delta as the difference between the PM₂.₅ predicted value and the counterfactual:

$$
\Delta = (PM₂.₅_{\text{predicted}}) - (\text{Counterfactual})
$$

Finally, we calculate the relative risk as the exponential of the product of beta and the mean of the yearly delta:

$$
\text{Relative Risk} = \exp \left( \beta \times \text{mean}(\Delta_{\text{yearly}}) \right)
$$



```{r, echo=FALSE}
rr <- targets::tar_read(calc_rr)
knitr::kable(rr,
             caption = "Relative Risk")

```

```{r, echo=FALSE}
paf <- targets::tar_read(calc_paf)
knitr::kable(paf,
             caption = "Population Attributable Fraction")

```

```{r, echo=FALSE}
an <- targets::tar_read(calc_an_all_ages)
knitr::kable(an, 
             caption = "Attributable Number" )
```

```{r, echo=FALSE}
an_sum <- targets::tar_read(calc_an_all_ages_sum)
knitr::kable(an_sum, 
             caption = "Attributable Number - Sum" )
```

```{r, echo=FALSE}
scale <- targets::tar_read(calc_scale_per_capita)
knitr::kable(scale, 
             caption = "Scaling - per capita" )
```

## Quality check

```{r, echo=FALSE}
qc_an <- targets::tar_read(an_all_ages_sum_qc)
knitr::kable(qc_an, 
             caption = "Quality Check - AN" )
```

### Fay H. Johnston 2020 
```{r, echo=FALSE}

j <- "figures_and_tables/qc_johnston.png"

# Include the plot in the report
knitr::include_graphics(j)
```
